The effects of binarity on planet occurrence rates measured by transit surveys

### 2018-01-30 paper/main.tex
incorporate JNW edits

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
This study develops simplified models of signal-to-noise limited transit 
surveys, in order to clarify the biases that stellar binarity introduces
to occurrence rate measurements.
We pay particular attention to dilution of observed planet radii, errors in 
star counts, and treatment of detection efficiencies.
Given a plausible set of assumptions, we derive a general formula 
for the apparent rate density, which depends on the true 
planet radius distribution, and also the relative numbers of planets 
orbiting secondaries, primaries, and single stars.
From this formalism, we suggest that for most of the observed radius 
distribution (apparent planet radii $>2r_\oplus$), ignoring binarity 
leads to errors in inferred occurrence rates at the $\lesssim 10\%$ level.
One corollary is that binarity seems unlikely to strongly influence any 
discrepancy between transit survey and RV survey hot Jupiter rates.
The errors created by ignoring binarity at smaller apparent radii might be 
larger: for instance, the absolute occurrence of Earth-sized planets could be 
overestimated by $\lesssim 50\%$.
Under our formalism, we also show that if high-resolution imaging reveals 
stellar neighbors of detected planets, the planet is usually more likely to 
orbit the primary. We explore how this statement depends on the
apparent radius of the detected planet.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A group of binarity-ignoring astronomers wants to measure the mean number of 
planets of a certain type per star of a certain type.
They perform a wide-field photometric search for planets that transit, and 
discover all for which
\begin{equation}
\left(\frac{{\rm S}}{{\rm N}}\right)
= ({\rm const}) \cdot \delta_{\rm obs} L_{\rm sys}^{1/2} d^{-1}
> \left(\frac{{\rm S}}{{\rm N}}\right)_{\rm floor}.
\label{eq:S_N_thresh}
\end{equation}
The observed transit depth per unit stellar flux $\delta_{\rm obs}$ 
constitutes the signal, ${\rm S}$.
The survey is shot-noise dominated, and the fractional photon noise 
${\rm N}$ scales as the inverse root of the photon flux received from any 
stellar system, ${\rm N} \propto (L_{\rm sys}/d^2)^{-1/2}$, for $L_{\rm sys}$ 
the system luminosity and $d$ its distance.
The solid angle coverage, telescope area, and survey duration are part of the 
``const'' term of Eq.~\ref{eq:S_N_thresh}. The transit duration also 
affects detectability, but we omit it in this work for brevity.

To compute an occurrence, at every planetary and stellar property the 
astronomers choose the stars around which the planets of interest appeared to 
be searchable.
Correcting for the geometric transit probability $p_{\rm tra}$, they report an 
apparent rate density,
\begin{equation}
\Gamma\a(\pp\a, \ps\a) = \frac{n_{\rm det}(\pp\a, \ps\a)}{N_{\rm s,a}(\pp\a, 
    \ps\a)} \times \frac{1}{p_{\rm tra}(\pp\a, \ps\a)}.
\label{eq:defn_apparent_rate_density}
\end{equation}
where $\pp\a$, $\ps\a$ are the apparent planetary and stellar parameters.
The quantity $N_{\rm s,a}(\pp\a, \ps\a)$ is the number of unresolved 
point-sources on the sky that appear to be searchable, and
$n_{\rm det}(\pp\a, \ps\a)$ is the number of detected planets, per unit 
$\pp\a$ and $\ps\a$.

There are many potential pitfalls.  Some genuine transit signals can be missed
by the detection pipeline.  Some apparent transit signals are spurious, from
noise fluctuations, failures of `detrending', or instrumental effects.  Stars
and planets can be misclassified due to statistical and systematic errors in
the measurements of their properties.  Poor angular resolution causes false
positives due to blends with background eclipsing binaries. {\it Et cetera}.

Here we focus on problems that arise from the fact that many stars exist in 
multiple star systems.
For simplicity, we consider only binaries, and we assume that they are all 
bound and spatially unresolved.

An immediate complication is that, due to dynamical stability or some 
aspect of planet formation, the occurrence rate of planets might differ 
between binary and single-star systems.
In an observed sample of stars that contains both singles and binaries, the 
detected planets would then be drawn from different intrinsic 
distributions for the occurrence rates in single star systems, 
about primaries, and about secondaries~\citep[{\it 
e.g.},][]{wang_occurrence_2015}.

Outside of astrophysical differences, every term in 
Eq.~\ref{eq:defn_apparent_rate_density} is observationally-biased.
There are errors in $n_{\rm det}(\pp\a,\ps\a)$ due to misclassification of 
both planet radii and stellar properties.
There are errors in $N_{{\rm s,a}}(\pp\a,\ps\a)$ because a 
searchable point on the sky might be two searchable stars.
Finally, there are errors in $p_{\rm tra}(\pp\a,\ps\a)$ because stars in 
binaries may have different masses and radii than assumed in the single-star 
case.

Correcting for binarity's observational biases is non-trivial.
For instance, in counting the number of searchable stars, even after realizing 
that binaries count as two stars, one must note that the multiplicity fraction 
of stars that are searchable for a given observed transit depth is greater 
than the multiplicity fraction in a volume limited sample.
This is the familiar Malmquist bias: at a fixed $\delta_{\rm obs}$, binaries 
are searchable out to larger distances than single stars because they are more 
luminous.
A separate effect is that the apparent radii assigned to planets in binaries 
will typically be smaller than the true radii.
Correcting for ``radius dilution'' requires knowing both the amount flux 
contributed by each source in a photometric aperture, and which star the 
planet orbits.

To gain intuition for the many observational biases at play,
we consider a set of idealized transit surveys.
First, we discuss the simplest possible cases (Sec.~\ref{sec:simplest}), in 
which all planets have identical properties, and all stars have identical 
properties except that some are in binary systems.
We generalize these thought-experiments in 
Sec.~\ref{sec:general_formula} to allow for realistic stellar populations with 
arbitrary occurrence rate distributions.
Sec.~\ref{sec:more_complicated} explores the implications for
(a) measurements of $\eta_\oplus$,
(b) the ``hot Jupiter rate discrepancy'', and 
(c) \citet{fulton_california-_2017}'s recently discovered ``valley''.
We discuss what our models suggest about the importance of binarity's biases 
in Sec.~\ref{sec:discussion}, and conclude in Sec.~\ref{sec:conclusion}.

@@

\section{The simplest possible transit survey models}
\label{sec:simplest}

\subsection{Twin binaries, identical stars, identical planets}
\label{sec:model_1}

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{
        Cartoon of the searchable volumes for single stars ({\it left}), and 
        twin binaries ({\it right}).
        This model assumes all stars have equal mass and luminosity, and 
        all planets have the same true radius $r=r\p$.
        {\it Top:} At an apparent radius $r\a=r\p$,
        the observer searched twin binaries out to a
        distance $\sqrt{2}\times$ that of single stars.
        Because of dilution, there are no planets in twin binaries with 
        $r\a=r\p$.
        {\it Middle:} At an apparent radius $r\a=r\p/\sqrt{2}$, the 
        maximum searchable distances are half those at $r\a=r\p$. 
        The only detected planets with $r\a=r\p/\sqrt{2}$ orbit twin binaries.
        {\it Bottom:} Planets with a true radius $r=r\p$ are searchable to 
        a maximum distance $d_0$ around singles, and $d_0/\sqrt{2}$ around 
        twin binaries.
    }
    \label{fig:model_1_volumes}
\end{figure}

Since the effects of binarity are most pronounced when the two components are 
similar, we begin by considering a universe in which all planets are 
identical, and all stars are identical except that some fraction of them exist 
in binaries. The goal of this thought experiment is to compare the occurrence 
inferred by observers ignoring binarity with the true occurrence in single 
star systems.

First, what are the possible apparent radii of detected planets?
If the true planet radius is $r=r\p$, then planets in singles will be 
detected with identical apparent radii: $r\a = r\p$.
Conversely, all detections in twin binaries will have apparent radii 
$r\a=r\p/\sqrt{2}$, because they are diluted.
For an unresolved point-source,
\begin{equation}
\delta_{\rm obs}
= \left(\frac{r\a}{R\a}\right)^2
= \left[{r\over R_{\rm host}}\right]^2\times {L_{\rm host} \over L_{\rm 
        sys}(M_1, q)},
\label{eq:delta_obs_general} 
\end{equation}
where $R_{\rm host}$ ($L_{\rm host}$) is the planet-host's stellar radius 
(luminosity), $R\a$ is the apparent stellar radius, $L_{\rm sys}=L_1+L_2 = 
L(M_1) + L(qM_1)$ is the system luminosity, and $q=M_2/M_1$ is the binary mass 
ratio.
We assume that the stellar radius and luminosity are uniquely related to the 
stellar mass.
The observers, with no a priori knowledge of the distance to a system, would 
measure a flux, and estimate a stellar mass $M\a=M_1$, a stellar 
radius $R\a = R(M\a)$, and an under-estimated system luminosity of $L_{\rm 
sys,a}(M\a) = L(M\a)$.

The planets will be detected if and only if their host stars are searchable.
For a fixed signal-to-noise floor, there is a maximum searchable 
distance~\citep{pepper_using_2003,pepper_searching_2005}.
From Eq.~\ref{eq:S_N_thresh}, this maximum searchable distance scales as
\begin{equation}
d(\delta_{\rm obs}, L_{\rm sys}) \propto \delta_{\rm obs} \cdot L_{\rm 
    sys}^{1/2}.
\end{equation}
Assuming that stars are uniformly distributed in space, the number 
of searchable stars $N\s$ is then proportional to
\begin{equation}
N\s(\delta_{\rm obs}, L_{\rm sys}) \propto n \delta_{\rm obs}^3 L_{\rm 
    sys}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n$ is the number per unit volume of {\it e.g.}, single star, or binary 
systems.
We neglect the dependence of $n$ on stellar type.

The searchable volumes for our thought experiment are illustrated 
in Fig.~\ref{fig:model_1_volumes}.
At an apparent radius equal to the true radius ($r\a=r\p$), single stars 
with $d<d_0$ are searchable.
Binaries with $r\a=r\p$ are also searchable, out to $\sqrt{2}d_0$. 
However, in this model no such systems exist; all planets in twin binaries 
have true radii $r=r\p$, and so are observed with apparent radii 
$r\a=r\p/\sqrt{2}$, out to $d_0/\sqrt{2}$.
At an apparent radius of $r\a=r\p/\sqrt{2}$, there could also be planets with 
true radii $r=r\p/\sqrt{2}$ orbiting singles~---~however none of these exist 
in this model.

How many planets are detected?
If we assume that there are $N\s^0(r\a)$ singles that are searchable for 
planets with $r\a$, and that there are $Z_0$ planets per single, then 
single stars contribute
\begin{equation}
n_{\rm det}^0(r\a) = N\s^0(r\a) Z_0 p_{\rm tra} \cdot \delta(r\a - r_p)
\end{equation}
planet detections (per unit $r\a$), where $\delta$ is the Dirac delta function.

We can write similar equations for the binaries.
First though, we define a ratio $\mu \equiv N\s^{\rm b}(r\a) / N\s^0(r\a)$ 
between the number of searchable binary systems and the number of searchable 
singles at any $r\a$.
This quantity is useful because although $N\s^{\rm b}$ and $N\s^0$ vary
as a function of the observed transit depth, $\mu$ remains fixed.
Applying this definition, at $r\a=r\p/\sqrt{2}$, some detections will come 
from primaries,
\begin{equation}
n_{\rm det}^1(r\a) =
\mu N\s^0(r\a) Z_1 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
and some from secondaries,
\begin{equation}
n_{\rm det}^2(r\a) = 
\mu N\s^0(r\a) Z_2 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
where $Z_1$ is the number of planets per primary, $Z_2$ is the number of 
planets per secondary, and $p_{\rm tra}$ is the same for identical stars.
In the case of twin binaries, Eq.~\ref{eq:N_searchable_prop} gives $\mu = 
2^{3/2} n_{\rm b}/n_{\rm s} = 2^{3/2} {\rm BF}/(1-{\rm BF})$, for ${\rm BF}$ 
the binary fraction\footnote{The binary fraction is the fraction of systems in 
a volume-limited sample that are binary. It is equivalent to the multiplicity 
fraction if there are no triple, quadruple, or higher order multiples.
}.

The last item we need to write down the apparent rate density $\Gamma\a(r\a)$ 
is the number of unresolved point-sources on sky that the observer thinks are 
searchable, $N_{\rm s,a}$.
Assuming that the apparent stellar radius is fixed, then given the 
apparent planet radius $r\a$,
\begin{equation}
N_{\rm s,a}(r\a) = N\s^0(r\a) \times (1 + \mu)
\label{eq:apparentlysearchable_to_searchable}
\end{equation}
relates the number of apparently searchable point-sources to the number of 
searchable singles.
From the discussion above, it should be clear that there are in fact
$N\s^0(r\a)\times (1 + 2\mu)$ searchable stars at any given $r\a$; the 
observers are under-counting.

Using Eq.~\ref{eq:defn_apparent_rate_density}, the apparent rate density 
computed without considering binarity is
\begin{align}
\Gamma\a(r\a) = 
\frac{1}{1+\mu} Z_0 \cdot
\delta(r\a-r\p)  +
\frac{\mu}{1+\mu} (Z_1 + Z_2) \cdot
\delta\left(r\a-\frac{r\p}{\sqrt{2}} \right).
\label{eq:model_1_apparent_rate_density}
\end{align}
The number of searchable singles at a given apparent radius, 
$N\s^0(r\a)$, makes no appearance.
The observer's errors in Eq.~\ref{eq:model_1_apparent_rate_density} are 
(A) under-counting the number of searchable stars at any apparent radius;
and
(B) inferring radii in binary systems that are all $\sqrt{2}$ too small.

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        Inferred occurrence rates over $0.01r_\oplus$ bins in planet 
        radius, for a model with identical stars, identical planets, and twin 
        binaries (same as Fig.~\ref{fig:model_1_volumes}).
        We assume a twin binary fraction of ${\rm BF}=0.1$, and that all 
        stars (single, primary, and secondary) host planets at equal 
        rates.
        If the true planet radius is $r\p$, all planets detected in binaries 
        will have apparent radii $r\a = r\p/\sqrt{2}$;
        Eq.~\ref{eq:model_1_apparent_rate_density} gives the normalizations.
        The rate and rate density are related by $\Lambda|_a^b = 
        \int_{a}^{b}\Gamma\,{\rm d}r$.
    }
    \label{fig:occ_rate_model_1}
\end{figure}

>>>>

\title{ The effects of binarity on planet occurrence rates measured by transit 
surveys}
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it Kepler}
and {\it TESS} missions, are usually conducted without knowing which stars
are actually binaries. Unresolved and unrecognized binary stars give rise to 
systematic errors in planet occurrence rates, including misclassified planets 
and miscounts in the number of searched stars. The individual errors can have 
different signs, making it difficult to anticipate the net effect on planet 
occurrence rates. Here we use simplified models of signal-to-noise limited 
transit surveys to try and clarify the situation. We derive a formula for the 
apparent occurrence rate density measured by an observer who falsely assumes 
all stars are single. The formula depends on the binary fraction; the 
mass function of the secondary stars; and the true occurrence of planets 
around primaries, secondaries, and single stars. It also takes into account 
the Malmquist bias by which binaries are over-represented in flux-limited 
samples. Application of the formula to an idealized {\it Kepler}-like survey 
shows that for planets larger than $2r_\oplus$, the net systematic error is of 
order 10\%. For smaller planets the errors are potentially larger: the 
occurrence of Earth-sized planets could be overestimated by as much as a 
factor of two. One consequence is that unrecognized binaries cannot account 
for the apparent discrepancy between the hot Jupiter occurrence rates measured 
by transit and RV surveys. We also demonstrate that if high-resolution imaging 
reveals a transit host star to be a binary, the planet is usually more likely 
to orbit the primary star than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

One of the goals of exoplanetary science is to establish how common,
or rare, are planets of various types.  Knowledge of planet occurrence
rates is helpful for inspiring and testing theories of planet
formation, designing the next generation of planet-finding surveys,
and simply satisfying our curiosity about other worlds.  One method
for measuring occurrence rates is to monitor the brightnesses of many
stars over a wide field, seeking evidence for planetary transits.
Measuring occurrence rates using this method was the highest priority
of the NASA {\it Kepler} mission.  Great strides have been made in the
analysis of {\it Kepler} data, including
progress toward the goal of determining the fraction of
Sun-like stars that harbor Earth-like 
planets~\citep{
    youdin_exoplanet_2011,petigura_prevalence_2013,dong_fast_2013,
    foreman-mackey_exoplanet_2014,burke_terrestrial_2015}

A lingering concern in occurrence rate studies is that in most cases,
investigators have assumed that all of the sources of light that were
monitored are single 
stars~\citep[\textit{e.g.},][]{
    howard_planet_2012,fressin_false_2013,foreman-mackey_exoplanet_2014,
    dressing_occurrence_2015,burke_terrestrial_2015}. In reality many of them 
are unresolved multiple-star systems, especially binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  
This ``flux dilution'' makes transit signals harder to detect and lowers the 
number of detections.  On the other hand, in a binary there two opportunities
to detect transiting planets, which could increase the overall number
of detections.

At the outset of this study it was not clear to us whether the neglect
of binaries was a serious issue at all, or even whether the net effect
of all the errors is positive or negative.  The goal of this study was
to clarify the various sources of error and provide a framework for
dealing with these issues.  In this spirit, most of our models are
idealized and analytic, and we do not attempt a detailed correction of
the results from {\it Kepler} or any other real transit survey.
However we can estimate the order of magnitude of the effects,
assuming the survey is signal-to-noise limited and given a model for
the true planet occurrence rate.

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then, in 
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets 
have identical properties, and all stars are identical
except that some fraction are in binary systems.  This simple 
model motivates the
derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more complicated but
nevertheless still analytic models.  In Section~\ref{sec:discussion}
we discuss the errors due to unrecognized binaries for some specific
cases of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot-Jupiter occurrence rates in different
surveys; and the shape of the ``evaporation valley'' in the radius
distribution of planets that was brought to light by
\citet{fulton_california-_2017}.  We review all the results in
Section~\ref{sec:conclusion}.

\section{Understanding the errors}
\label{sec:concept}

Imagine a group of astronomers that wants to measure the mean number of
planets of a certain type per star of a certain type.  They perform a
wide-field photometric search for planets that transit, and discover
all for which
\begin{equation}
\frac{\delta_{\rm obs}}{\sigma}
  = ({\rm constant}) \cdot \delta_{\rm obs} L_{\rm sys}^{1/2} d^{-1}
> \left(\frac{{\rm S}}{{\rm N}}\right)_{\rm min}.
\label{eq:S_N_thresh}
\end{equation}
Here the signal S is the observed transit depth $\delta_{\rm obs}$, in
units of relative flux.
%LB: we don't introduce r or R because they are system-specific.
%For single stars, $\delta_{\rm obs} \approx (r/R)^2$; for binaries,
%flux dilution $\mathcal{D}$ lowers the observed depth to
%$\delta_{\rm obs} \approx \mathcal{D} (r/R)^2$. 
The noise N is the uncertainty in the
determination of the flux ratio of the star inside and outside of
transits.  Throughout this paper we will assume the photometric
uncertainty to scale as the inverse square root of the number of
photons collected from the source, {\it i.e.}, the dependence on
stellar properties is $\sigma \propto (L_{\rm sys}/d^2)^{-1/2}$, where
$L_{\rm sys}$ is the luminosity of the stellar system and $d$ is its
distance from Earth.
% The solid angle coverage, telescope area, and
% survey duration are in the ``constant'' term of
% Equation~\ref{eq:S_N_thresh}.  The transit duration also affects
% detectability, but we omit it in this work for brevity.

The astronomers then assume that all stars are single.  In particular
they do not have accurate enough parallaxes to tell that some of the
stars are apparently overluminous. 
They want to compute the number of planets with size $r$ per star of mass $M$.
If the total number of such stars for which the planets could be detected
($N_\star$) and the number of detected planets ($N_{\rm det}$) are both 
large, then the astronomers' estimate for the occurrence rate ($\Lambda$) is 
simply
\begin{equation}
\Lambda(r,M) = \frac{N_{\rm det}(r,M)}{N_\star(r,M)}
                    \times \frac{1}{p_{\rm tra}(r,M)},
\label{eq:occ_rate_simple}
\end{equation}
where $p_{\rm tra}$ is the geometric transit probability, needed to
account for the fact that most planetary orbits are not aligned close
enough with our line of sight to produce transits.

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic errors in the catalogued properties of the stars.  Some
apparent transit signals are spurious, arising from noise fluctuations
or failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

We will focus exclusively on the subset of problems that arise from
the fact that many stars exist in gravitationally bound binary
systems.  Even with this narrow focus, there are numerous sources of
error.  All three of the quantities in
Equation~\ref{eq:occ_rate_simple} are subject to
observational bias:
\begin{enumerate}
%    
    \item The number of detections, $N_{\det}$, is actually the number of 
    detected planets that {\it appear} to have size $r$, orbiting stars that 
    {\it appear} to have mass $M$.
    Whenever the planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size will be misclassified because of the reduction
        in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified because its 
        light is combined with a second star of a different spectral type.
%        
    \end{itemize}
%    
    \item The apparent number of stars that were searched, $N_\star$,
    is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of the
        secondary stars that were searched for transiting planets.
%        
        \item toward higher values, because some of the stars that
        appeared to be suitably small and bright to detect the desired
        type of planet are in fact binaries for which the amplitude of
        the photometric signal would be reduced to an undetectable
        level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$, which scales with the
    stellar mean density as $\rho^{-1/3}$, is biased because the
    planet-hosting star could be misclassified.
%    
\end{enumerate}

Two other problems that may arise are not represented in
Equation~\ref{eq:occ_rate_simple}.  One is astrophysical:
the true occurrence rate of the desired type of planet may depend on
whether the host is a single star, the primary star of a binary, or
the secondary star of a binary.  Such differences could be caused by
the requirement for long-term dynamical stability, or differences in
the planet formation process.  When the search sample includes both
singles and binaries, the detected planets are thereby drawn from
different occurrence distributions~\citep[see,
{\it e.g.},][]{wang_occurrence_2015}.

The other problem is an observational effect.  Even after the
astronomers learn about binaries and attempt to correct for their
presence, they must realize that the ratio of binaries to single stars
in the sample will differ from that in a volume-limited sample due to
a type of Malmquist bias.  Binaries have a higher total luminosity
than either the primary or secondary star would have on its own.  This
means that the binaries are searchable for transit signals of a given
amplitude at greater distances from the Earth.  The binaries in a
sample of apparently searchable stars are therefore drawn from a
larger search volume than the single stars.

% JNW: Let's remember, when discussing Kepler, to point out that it's
% not really a SNR or flux limited survey, because of the
% preselection. This would need to be dealt with in a realistic
% treatment.

% JNW: Let's also remember to say somewhere that we do not deal with
% period distributions none of these effects mess up the period.

Planet occurrence rates are often presented as two-dimensional density
functions in the space of planet radius and orbital period.
Thankfully none of the errors we have enumerated lead to errors in the
period distribution.  When more than one transit is detected (as is
usually required by the surveyors), the orbital periods can be
measured without ambiguity regardless of whether the host star is
single or one member of a binary.  Hence in what follows, we focus on
the radius distribution and assume the periods are measured without
significant uncertainty.

Given all of the confusing and opposing sources of error, we will
proceed in stages. We will start from a barebones model in which
everything can be written down on the back of a napkin, and build up
to an analytic model allowing for generality in the distribution of
the binaries and the planets they host.  We consider only the
corrections to the occurrence as a function of planetary radius, and
not the orbital period (the other key variable in transit studies),
because when multiple transits are detected the period can be directly
measured regardless of whether the host star is single or part of a
binary.

@@

\section{Simple models}
\label{sec:simplest}

\subsection{One type of star, one type of planet}
\label{sec:model_1}

Since the effects of binarity are most pronounced when the two
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Some fraction of them exist in twin binaries.
All planets have size $r\p$.
Our goal is to compare the occurrence inferred by observers who
ignore binarity with the true occurrence in single star systems.

To clarify our terminology, we define an ``apparent planet radius'', $r\a$, as 
the radius that one would infer from a transit signal under the assumption 
that the host star is single.
The observed transit depth is given by the ratio of the apparent planet
radius and the apparent stellar radius, $R\a$:
\begin{equation}
\delta_{\rm obs}
= \left(\frac{r\a}{R\a}\right)^2
= \left[{r\over R_{\rm host}}\right]^2\times {L_{\rm host} \over L_{\rm 
        sys}},
\label{eq:delta_obs_general} 
\end{equation}
where $r$ is the planet radius, $R_{\rm host}$ ($L_{\rm host}$) is the 
host star's radius (luminosity), and $L_{\rm sys}$ is the 
entire system's luminosity.
For the example at hand, $R_{\rm host} = R$, since all stars have the same 
size. In single star systems we have $r=r\a=r\p$. In twin binaries, 
$r\a=r\p/\sqrt{2}$, since $L_{\rm sys}=2L$.

We can now understand that the observers are unwittingly
reporting an {\it apparent} occurrence rate $\Lambda\a(r\a)$ as
a function of apparent radius.
The apparent occurrence rate is the same as the {\it true} occurrence rate for 
transit signals of 
amplitude $\delta_{\rm obs} = (r\a/R\a)^2$, \textit{i.e.}, the number of 
detections ($N_{\rm det}$) of signals of amplitude $\delta_{\rm obs}$, divided 
by the number of sources $N_\star$ from which such a signal could be detected.

Finally, it is helpful to introduce an apparent rate density 
$\Gamma\a$, the average number of planets per star per unit apparent planet 
radius. Analogous to Equation~\ref{eq:occ_rate_simple},  
\begin{equation}
\Gamma\a(r\a) = \frac{n_{\rm det}(r\a)}{N_\star(r\a)}
\times \frac{1}{p_{\rm tra}(r\a)},
\label{eq:apparent_rate_density}
\end{equation}
where $n_{\rm det}={\rm d}N_{\rm det}/{\rm d}r\a$ is the number  
of detected planets, per unit $r\a$, with apparent radius $r\a$.
In Equation~\ref{eq:apparent_rate_density}, we neglect the dependence on 
stellar mass since all stars in this model have the same properties.

To compute the number of detected planets, we consider the volumes within 
which a given signal could be detected. At a fixed signal-to-noise floor, 
there is a maximum searchable 
distance~\citep{pepper_using_2003,pepper_searching_2005}.
From Equation~\ref{eq:S_N_thresh}, this maximum searchable distance scales as
\begin{equation}
d \propto \delta_{\rm obs} \cdot L_{\rm 
    sys}^{1/2}.
\end{equation}
Assuming that stars are uniformly distributed in space, the number 
of searchable stars $N_\star$ is then proportional to
\begin{equation}
N_\star \propto n \delta_{\rm obs}^3 L_{\rm 
    sys}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n$ is the number per unit volume of {\it e.g.}, single star, or binary 
systems.
We neglect the dependence of $n$ on stellar type.

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{
        Searchable volumes for single stars ({\it left}) and 
        twin binaries ({\it right}), assuming that all stars have the 
        same mass, radius, and luminosity, and that all planets have the same 
        radius $r=r\p$.
        {\it Top:} At an apparent radius $r\a=r\p$,
        the observer searched twin binaries out to a
        distance $\sqrt{2}\times$ that of single stars.
        Because of dilution, there are no planets in twin binaries with 
        $r\a=r\p$.
        {\it Middle:} At $r\a=r\p/\sqrt{2}$, the 
        maximum searchable distances are half those at $r\a=r\p$. 
        The only detected planets with $r\a=r\p/\sqrt{2}$ orbit twin binaries.
        {\it Bottom:} Planets with a true radius $r=r\p$ are searchable to 
        a maximum distance $d_0$ around singles, and $d_0/\sqrt{2}$ around 
        twin binaries.
    }
    \label{fig:model_1_volumes}
\end{figure}

The searchable volumes for this example are illustrated in 
Figure~\ref{fig:model_1_volumes}.
At an apparent radius equal to the true radius ($r\a=r\p$), single stars 
with $d<d_0$ are searchable.
Binaries with $r\a=r\p$ are also searchable, out to $\sqrt{2}d_0$. 
However, in this example no such systems exist; all planets in twin binaries 
have true radii $r=r\p$, and so are observed with apparent radii 
$r\a=r\p/\sqrt{2}$, out to $d_0/\sqrt{2}$.
At an apparent radius of $r\a=r\p/\sqrt{2}$, there could also be planets with 
true radii $r=r\p/\sqrt{2}$ orbiting singles, but there are no such systems 
in this model.

How many planets are detected?
If we assume that there are $N_\star^0(r\a)$ singles that are searchable for 
planets with $r\a$, and that there are $Z_0$ planets per single, then 
single stars contribute
\begin{equation}
n_{\rm det}^0(r\a) = N_\star^0(r\a) Z_0 p_{\rm tra} \cdot \delta(r\a - r_p)
\end{equation}
planet detections (per unit $r\a$), for $\delta$ the Dirac delta function.

We can write similar equations for the binaries.
First though, we define a ratio $\mu \equiv N_\star^{\rm b}(r\a) / 
N_\star^0(r\a)$ 
between the number of searchable binary systems and the number of searchable 
singles at any $r\a$.
This quantity is useful because although $N_\star^{\rm b}$ and $N_\star^0$ vary
as a function of the observed transit depth, $\mu$ remains fixed.
Applying this definition, at $r\a=r\p/\sqrt{2}$, some detections will come 
from primaries,
\begin{equation}
n_{\rm det}^1(r\a) =
\mu N_\star^0(r\a) Z_1 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
and some from secondaries,
\begin{equation}
n_{\rm det}^2(r\a) = 
\mu N_\star^0(r\a) Z_2 p_{\rm tra}
\cdot \delta\left(r\a - \frac{r\p}{\sqrt{2}}\right),
\end{equation}
where $Z_1$ is the number of planets per primary, $Z_2$ is the number of 
planets per secondary, and $p_{\rm tra}$ is the same for identical stars.
For twin binaries, Equation~\ref{eq:N_searchable_prop} gives $\mu = 
2^{3/2} n_{\rm b}/n_{\rm s}$, where $n_{\rm b}$ is the number density of 
binaries in a volume-limited sample.

The last item we need to write down the apparent rate density $\Gamma\a(r\a)$ 
is the number of unresolved point-sources on sky that the observer thinks are 
searchable, $N_\star$.
From the definition of $\mu$,
\begin{equation}
N_\star(r\a) = N_\star^0(r\a) \times (1 + \mu)
\label{eq:apparentlysearchable_to_searchable}
\end{equation}
relates the number of apparently searchable point-sources to the number of 
searchable singles.
From the discussion above, it should be clear that there are in fact
$N_\star^0(r\a)\times (1 + 2\mu)$ searchable stars at any given $r\a$; the 
observers are under-counting.

Using Equation~\ref{eq:apparent_rate_density}, the apparent rate density 
computed when ignoring binarity is
\begin{align}
\Gamma\a(r\a) = 
\frac{1}{1+\mu} Z_0 \cdot
\delta(r\a-r\p)  +
\frac{\mu}{1+\mu} (Z_1 + Z_2) \cdot
\delta\left(r\a-\frac{r\p}{\sqrt{2}} \right).
\label{eq:model_1_apparent_rate_density}
\end{align}
The number of searchable singles at a given apparent radius, 
$N_\star^0(r\a)$, makes no appearance.
The observer has erred in Equation~\ref{eq:model_1_apparent_rate_density} by 
under-counting the number of searchable stars at any apparent radius, and by
inferring incorrect planet radii in binary systems.
